Correcting K-correction: dark energy is based on a math error from 1930

### 2025-01-05 paper/dark_energy_corrections.tex

\author[0000-0001-6450-3262]{Logan P. Evans}

@@

\begin{figure}
  \includegraphics[width=\linewidth]{mu_distance_vs_redshift.png}
  \caption{The relationship between distance and redshift for two treatments of
  magnitude data. The displayed points are roughly a third of the values in the
  full SnIa dataset published in \citet{abbott2024}, selected evenly to aid
  visibility. The $k(z) = 1$ treatment, which represents the uncorrected
  values, is clearly non-linear while the $k(z) = 1 + z$ treatment which fixes
  the K-correction redshift error, appears to be linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

@@

\begin{figure}
  \includegraphics[width=.75\linewidth, height=0.75\vsize]{k_equation_flow.png}
  \caption{The ratio $K'_{xy}$ is the crux of the K-correction algorithm. We
  are able to observe a SnIa in the redshifted observation frame, but we want
  to compute the magnitude in the rest frame as if no redshift has occurred.
  This computation starts with $F(\lambda$, the spectral energy density that
  describes the expected luminosity in the rest frame. Using the SED, we can
  compute the photon density $F'(\lambda)$ for the rest frame. We show the
  process of computing $R'(\lambda$, the photon density in the observation
  frame, in three phases starting at the top right of the figure. First we
  stretch the photon density, but this step inflates counts. In the second
  step, we correct for the inflated counts by multiplying by $1/(1+z)$;
  at this point, the total count equals the total count of $F'(\lambda)$.
  Finally, in the third step we finish calculating the observation frame photon
  density $R'(\lambda)$ by accounting for time dilation, which also reduces
  photon counts by a factor of $1/(1+z)$.  Given the rest frame and
  observation frame photon densities, we can compute the expected ratio of
  photons by totalling the area under the curves in the two panels at the
  bottom of the graph. This example does not show the process of converting
  photon flux to magnitude, nor does it address the zero-point correction.
  }
  \label{fig:k-example}
\end{figure}

@@

@@

@@

The value estimated here, $H_0 \sim 65.94 \pm 1.29$ is highly suspicious
because of the flipped zero-point corrections, but it is consistent with
estimates of $H_0$ that are based on the CMB. \citet{planck2020} published the
CMB value of $H_0 \sim 67.27 \pm 0.6$, and the one sigma error bars overlap. It
is interesting to note that restricting the Dark Energy Survey observations of
SnIa \citep{vincenzi2024} to only those with $z < 0.1$, which presumably is a
small enough redshift that all observations will have been made with the same
filter, gives the estimate $H_0 \sim 67.8 \pm 1.86$.

\begin{figure}
  \includegraphics[width=\linewidth]{velocity_vs_distance.png}
  \caption{The relationship between expansion velocity and distance. The slope of this graph demonstrates the Hubble constant.
  }
\label{fig:expansion}
\end{figure}

\begin{figure}
  \includegraphics[width=\linewidth]{bootstrapped_H0.png}
  \caption{A histogram of $100{\small,}000$ bootstrap trials measuring the Hubble constant
  $H_0$. Each trial samples an absolute magnitude $M \sim Norm(-19.2334,
  0.0404)$ for SnIa based on data published by \citet{camarena2020}. It then samples,
  with replacement, a population of supernovae from the dataset published by
  \citet{abbott2024}. Finally, it uses the non-parametric linear regression
  technique described by \citet{siegel1982}. The result of the bootstrap is $H_0
  \sim Norm(65.94, 1.29^2)$.
  }
\label{fig:H0bootstrap}
\end{figure}

>>>>

\author[0000-0001-6450-3262]{Logan P. Evans}
\noaffiliation

@@

\begin{figure}
  \plotone{lum_distance_vs_redshift.pdf}
  \caption{The relationship between luminosity distance and redshift for two
  treatments of magnitude data. The displayed points are roughly a third of the
  values in the full SnIa dataset published in \citet{abbott2024}, selected
  evenly to aid visibility. The $k(z) = 1$ treatment, which represents the
  uncorrected values, is clearly non-linear while the $k(z) = 1 + z$ treatment
  which fixes the K-correction redshift error, appears to be linear.}
  \label{fig:mu_distance_vs_redshift}
\end{figure}

@@

@@

\begin{figure}
  \plotone{k_equation_flow.pdf}
  \caption{The ratio $K'_{xy}$ is the crux of the K-correction algorithm. We
  are able to observe a SnIa in the redshifted observation frame, but we want
  to compute the magnitude in the rest frame as if no redshift has occurred.
  This computation starts with $F(\lambda$, the spectral energy density, which
  we then convert to $F'(\lambda)$, the photon density for the rest frame. We
  show the process of computing $R'(\lambda$, the photon density in the
  observation frame, in three phases starting at the top right of the figure.
  First we stretch the photon density, but this step inflates counts. In the
  second step, we correct for the inflated counts by multiplying by $1/(1+z)$;
  at this point, the total count equals the total count of $F'(\lambda)$.
  Finally, in the third step we finish calculating the observation frame photon
  density $R'(\lambda)$ by accounting for time dilation, which also reduces
  photon counts by a factor of $1/(1+z)$. We can compute the expected ratio of
  photons by totalling the area under the curves in the bottom two panels. This
  example does not show the process of converting photon flux to magnitude, nor
  does it address the zero-point correction.
  }
  \label{fig:k-example}
\end{figure}

@@

\begin{figure}
  \plotone{velocity_vs_distance.pdf}
  \caption{The relationship between expansion velocity and distance. The slope
  of this graph demonstrates the Hubble constant. The data for these SnIa
  observations comes from the Dark Energy Spectroscopic Instrument
  \citep{vincenzi2024}, but the magnitudes are corrected to account for time
  dilation by adding $-2.5 log(1 + z)$.
  }
\label{fig:expansion}
\end{figure}

@@

\begin{figure}
  \plotone{bootstrapped_H0.pdf}
  \caption{A histogram of $10{,}000$ bootstrap trials measuring the Hubble constant
  $H_0$. Each trial samples an absolute magnitude $M \sim Norm(-19.2334,
  0.0404)$ for SnIa based on data published by \citet{camarena2020}. It then samples,
  with replacement, a population of supernovae from the dataset published by
  \citet{abbott2024}. Finally, it uses the non-parametric linear regression
  technique described by \citet{siegel1982}. The result of the bootstrap is $H_0
  \sim Norm(65.94, 1.30^2)$.
  }
\label{fig:H0bootstrap}
\end{figure}

The value estimated here, $H_0 \sim 65.94 \pm 1.30$ is highly suspicious
because of the flipped zero-point corrections, but it is consistent with
estimates of $H_0$ that are based on the CMB. \citet{planck2020} published the
CMB value of $H_0 \sim 67.27 \pm 0.6$, and the one sigma error bars overlap. It
is interesting to note that restricting the Dark Energy Survey observations of
SnIa \citep{vincenzi2024} to only those with $z < 0.1$, which presumably is a
small enough redshift that all observations will have been made with the same
filter, gives the estimate $H_0 \sim 67.8 \pm 1.86$.
